Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems

### 2016-09-07 talk/talk.tex
More animation

\begin{frame}{Connectedness as a Constraint (on Variables)}
    \begin{itemize}
        \item Once we have assigned one variable to a non-$\bot$ value, traverse the variable graph,
            ignoring any variables assigned to $\bot$. Assign any unvisited variables to $\bot$.
        \item Repeat whenever any variable is assigned to $\bot$ during search.
    \end{itemize}
\end{frame}

\begin{frame}{Connectedness as a Constraint (on Values)}
    \begin{itemize}
        \item Also, traverse the value graph, ignoring any values not present in any domain. Remove
            any unvisited values from every domain.
        \item Repeat whenever a value is removed from every domain during search.
    \end{itemize}
\end{frame}

\begin{frame}{Connectedness via Restricted Branching}
    \begin{itemize}
        \item When selecting a variable for branching (except for the first variable), only select
            from variables which are adjacent to at least one variable already assigned to
            non-$\bot$.
        \item If no such variable exists, assign every remaining domain to $\bot$.
    \end{itemize}
\end{frame}

@@

\begin{frame}{Connectedness in Clique-Like Algorithms}
    \begin{itemize}
        \item Connectedness doesn't decompose cleanly into binary constraints, so we can't encode it
            as microstructure.
        \item On the other hand, we \emph{can} use restricted branching, if we give more information
            to the clique algorithm.
            \begin{itemize}
                \item Integrating this with the bound and ordering heuristic used in modern clique
                    algorithms is horribly technical and was far harder than we make it look in the
                    paper (honest!).
            \end{itemize}
    \end{itemize}
\end{frame}

>>>>

\begin{frame}{Connectedness as a Constraint (on Variables)}
    \begin{itemize}
        \item We can do better than just verifying connectedness:
            \begin{itemize}
                \item Once we have assigned one variable to a non-$\bot$ value, traverse the variable graph,
                    ignoring any variables assigned to $\bot$. Assign any unvisited variables to $\bot$.
                \item Repeat whenever any variable is assigned to $\bot$ during search.
            \end{itemize}
    \end{itemize}

    \vspace{1em}

    \begin{center}
    \begin{tikzpicture}[scale=0.75]%{{{
        \begin{scope}
            \node <1>   [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Ma) at (0, 0) {\phantom{0}};
            \node <2->  [draw, circle, fill=uofgcobalt, inner sep=2.5pt, font=\normalsize] (Ma) at (0, 0) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Mb) at (2, 0) {\phantom{0}};
            \node <1-2> [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Mc) at (2, -2) {\phantom{0}};
            \node <3->  [draw, circle, fill=white, draw=uofgsandstone!60, inner sep=2.5pt, font=\normalsize] (Mc) at (2, -2) {\phantom{0}};
            \node <1-3> [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Md) at (4, 0) {\phantom{0}};
            \node <4->  [draw, circle, fill=white, draw=uofgsandstone!60, inner sep=2.5pt, font=\normalsize] (Md) at (4, 0) {\phantom{0}};
            \node <1-4> [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Me) at (4, -2) {\phantom{0}};
            \node <1-4> [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Mf) at (6, -2) {\phantom{0}};
            \node <5->  [draw, circle, fill=white, draw=uofgsandstone!60, inner sep=2.5pt, font=\normalsize] (Me) at (4, -2) {\phantom{0}};
            \node <5->  [draw, circle, fill=white, draw=uofgsandstone!60, inner sep=2.5pt, font=\normalsize] (Mf) at (6, -2) {\phantom{0}};

            \draw <1-2> [very thick] (Mc) -- (Me);
            \draw <1-2> [very thick] (Ma) -- (Mc);
            \draw <1-2> [very thick] (Mb) -- (Mc);

            \draw <3->  [color=uofgsandstone!60] (Mc) -- (Me);
            \draw <3->  [color=uofgsandstone!60] (Ma) -- (Mc);
            \draw <3->  [color=uofgsandstone!60] (Mb) -- (Mc);

            \draw <1-3> [very thick] (Mb) -- (Md);
            \draw <1-3> [very thick] (Md) -- (Me);
            \draw <4->  [color=uofgsandstone!60] (Mb) -- (Md);
            \draw <4->  [color=uofgsandstone!60] (Md) -- (Me);

            \draw <1-4> [very thick] (Me) -- (Mf);
            \draw <5->  [color=uofgsandstone!60] (Me) -- (Mf);

            \draw [very thick] (Ma) -- (Mb);
        \end{scope}
    \end{tikzpicture}
    \end{center}
\end{frame}

\begin{frame}{Connectedness as a Constraint (on Values)}
    \begin{itemize}
        \item We can do something similar on the dual viewpoint:
            \begin{itemize}
                \item Also, traverse the value graph, ignoring any values not present in any domain. Remove
                    any unvisited values from every domain.
                \item Repeat whenever a value is removed from every domain during search.
        \end{itemize}
    \end{itemize}

    \vspace{1em}

    \begin{center}
    \begin{tikzpicture}[scale=0.75]%{{{
        \begin{scope}
            \node <1>   [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Ma) at (0, 0) {\phantom{0}};
            \node <2->  [draw, circle, fill=uofgcobalt, inner sep=2.5pt, font=\normalsize] (Ma) at (0, 0) {\phantom{0}};
            \node <1->  [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Mb) at (1, -1) {\phantom{0}};
            \node <1->  [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Mc) at (1, 1) {\phantom{0}};
            \node <1-2>  [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Md) at (2, 0) {\phantom{0}};
            \node <3->  [draw, circle, fill=white, draw=uofgsandstone!60, inner sep=2.5pt, font=\normalsize] (Md) at (2, 0) {\phantom{0}};
            \node <1-3>  [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Me) at (3, -1) {\phantom{0}};
            \node <1-3>  [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Mf) at (3, 1) {\phantom{0}};
            \node <1-3>  [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Mg) at (4, 0) {\phantom{0}};
            \node <4->  [draw, circle, fill=white, draw=uofgsandstone!60, inner sep=2.5pt, font=\normalsize] (Me) at (3, -1) {\phantom{0}};
            \node <4->  [draw, circle, fill=white, draw=uofgsandstone!60, inner sep=2.5pt, font=\normalsize] (Mf) at (3, 1) {\phantom{0}};
            \node <4->  [draw, circle, fill=white, draw=uofgsandstone!60, inner sep=2.5pt, font=\normalsize] (Mg) at (4, 0) {\phantom{0}};

            \draw <1-2> [very thick] (Mb) -- (Md);
            \draw <1-2> [very thick] (Mc) -- (Md);
            \draw <1-2> [very thick] (Md) -- (Me);
            \draw <1-2> [very thick] (Md) -- (Mf);
            \draw <3->  [draw=uofgsandstone!60] (Mb) -- (Md);
            \draw <3->  [draw=uofgsandstone!60] (Mc) -- (Md);
            \draw <3->  [draw=uofgsandstone!60] (Md) -- (Me);
            \draw <3->  [draw=uofgsandstone!60] (Md) -- (Mf);

            \draw <1-3> [very thick] (Me) -- (Mg);
            \draw <1-3> [very thick] (Mf) -- (Mg);
            \draw <4->  [draw=uofgsandstone!60] (Me) -- (Mg);
            \draw <4->  [draw=uofgsandstone!60] (Mf) -- (Mg);

            \draw [very thick] (Ma) -- (Mb);
            \draw [very thick] (Ma) -- (Mc);
        \end{scope}
    \end{tikzpicture}
    \end{center}
\end{frame}

\begin{frame}{Connectedness via Restricted Branching}
    \begin{itemize}
        \item The following heretical approach is due to Vismara and Valery.
        \item When selecting a variable for branching (except for the first variable), only select
            from variables which are adjacent to at least one variable already assigned to
                    non-$\bot$.
        \item If no such variable exists, assign every remaining domain to $\bot$.
    \end{itemize}

    \vspace{1em}

    \begin{center}
    \begin{tikzpicture}[scale=1.25]%{{{
        \begin{scope}
            \node <1>   [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Ma) at (0, 0) {\phantom{0}};
            \node <2->  [draw, circle, fill=uofgcobalt, inner sep=2.5pt, font=\normalsize] (Ma) at (0, 0) {\phantom{0}};
            \node <1-3> [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Mb) at (1, 0) {\phantom{0}};
            \node <4->  [draw, circle, fill=uofgpumpkin, inner sep=2.5pt, font=\normalsize] (Mb) at (1, 0) {\phantom{0}};
            \node <1-6>  [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Mc) at (1, -1) {\phantom{0}};
            \node <7->  [draw, circle, fill=white, draw=uofgsandstone!60, inner sep=2.5pt, font=\normalsize] (Mc) at (1, -1) {\phantom{0}};
            \node <1-2> [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Md) at (2, 0) {\phantom{0}};
            \node <1-2> [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Me) at (2, -1) {\phantom{0}};
            \node <3-4>  [draw, circle, dotted, fill=white, inner sep=2.5pt, font=\normalsize] (Md) at (2, 0) {\phantom{0}};
            \node <5>  [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Md) at (2, 0) {\phantom{0}};
            \node <6->  [draw, circle, fill=uofglawn, inner sep=2.5pt, font=\normalsize] (Md) at (2, 0) {\phantom{0}};
            \node <3-7>  [draw, circle, dotted, fill=white, inner sep=2.5pt, font=\normalsize] (Me) at (2, -1) {\phantom{0}};
            \node <8->  [draw, circle, draw=uofgsandstone!60, fill=white, inner sep=2.5pt, font=\normalsize] (Me) at (2, -1) {\phantom{0}};
            \node <1-2> [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Mf) at (3, 0) {\phantom{0}};
            \node <3-7>  [draw, circle, dotted, fill=white, inner sep=2.5pt, font=\normalsize] (Mf) at (3, 0) {\phantom{0}};
            \node <8-> [draw, circle, fill=white, draw=uofgsandstone!60, inner sep=2.5pt, font=\normalsize] (Mf) at (3, 0) {\phantom{0}};
            \node <1-2> [draw, circle, fill=white, inner sep=2.5pt, font=\normalsize] (Mg) at (3, -1) {\phantom{0}};
            \node <3-7>  [draw, circle, dotted, fill=white, inner sep=2.5pt, font=\normalsize] (Mg) at (3, -1) {\phantom{0}};
            \node <8->  [draw, circle, draw=uofgsandstone!60, fill=white, inner sep=2.5pt, font=\normalsize] (Mg) at (3, -1) {\phantom{0}};

            \draw [very thick] (Ma) -- (Mb);
            \draw [very thick] (Mb) -- (Md);
            \draw <1-6> [very thick] (Mc) -- (Me);
            \draw <1-6> [very thick] (Ma) -- (Mc);
            \draw <1-6> [very thick] (Mb) -- (Mc);
            \draw <7-> [draw=uofgsandstone!60] (Mc) -- (Me);
            \draw <7-> [draw=uofgsandstone!60] (Ma) -- (Mc);
            \draw <7-> [draw=uofgsandstone!60] (Mb) -- (Mc);
            \draw <1-7> [very thick] (Mf) -- (Mg);
            \draw <8-> [draw=uofgsandstone!60] (Mf) -- (Mg);
        \end{scope}
    \end{tikzpicture}
    \end{center}
\end{frame}

@@

\begin{frame}{Connectedness in Clique-Like Algorithms?}
    \begin{itemize}
        \item Connectedness doesn't decompose cleanly into binary constraints,
            so we can't encode it as microstructure.
        \item Modern maximum clique algorithms use a special bound function
            which doesn't allow us to branch arbitrarily. Adapting the
            restricted branching rule to work with this would be horribly
            complicated and no-one in their right mind would even attempt it.
            \begin{itemize}
                \item <2-> But we did it anyway. See the paper for how.
            \end{itemize}
    \end{itemize}
\end{frame}
